Dynamical dark energy constraints via clustering shells of SDSS-III BOSS galaxies

### 2017-07-13 Draft/Draft0.3.tex
minor changes

\author{ Xiao-Dong Li,$^{1,\dagger}$
Cristiano G. Sabiu,$^{2}$
Changbom Park,$^{1}$
Yuting Wang,$^{3,\star}$
Gong-bo Zhao,$^{3}$
Hyunbae Park,$^{2}$
Arman Shafieloo,$^{2}$
Juhan Kim,$^{4,1}$
and Sungwook E. Hong$^{2}$}

@@

\begin{abstract}
We use the anisotropic {\it clustering shells} of 1,133,326 galaxies from the Sloan Digital Sky Survey Data Release 
12 spanning the redshift range $0.15<z<0.69$ to place constraints on the expansion history of the Universe. 
Concentrating on the Chevallier-Polarski-Linder parametrization of dark energy $w=w_0+w_a\frac{z}{1+z}$,
while improving upon earlier works \citep{Li2016}, 
we present a novel use of the Alcock-Paczynski (AP) test via the small scales clustering difference between redshift bins. 
This differential measure appears to be a robust statistical tool which mitigates many systematic effects, 
allowing for competitive and unbiased constrains. 
A combination of CMB+BAO+SNIa+$H_0$+AP yields
$\Omega_m = 0.301 \pm 0.008,\ w_0 = -1.042 \pm 0.067,\ w_a = -0.07 \pm 0.29$ (68.3\% CL).
Adding the AP method reduces the $w_0-w_a$ contour area by 50\%.
The method can be combined with the standard BAO method, 
increasing the science output of ongoing and future galaxy surveys by a factor of 2.
\end{abstract}

@@

%\section{Introduction}
{\it Introduction.}---
The  origin  of  the  accelerating  expansion  of  the  universe  is  one of  the  most  salient  questions  in  contemporary cosmology. 
Proposed mechanisms include  evolving  scalar  fields  remnant  from  the  big  bang, 
modifications to gravity,  and a  non-zero  vacuum energy \cite[see][for a comprehensive review]{2012IJMPD..2130002Y}. 
Considering the wealth of theoretical explanations, it is crucial to obtain precise and unbiased measurements of the expansion history of the Universe allowing us to 
differentiate between competing dark energy models.

@@

In an effort to mitigate the RSD effect, \cite{Li2014} proposed utilizing the {\it redshift dependence} of the AP effect. 
The clustering anisotropies produced by are, although large, close to uniform in magnitude over a wide range in redshift. 
However, if cosmological parameters are incorrectly chosen resulting in an AP effect, 
the anisotropy in the clustering signal has a clear redshift dependence. 
In a later work, \cite{Li2015} developed an AP methodology 
that utilises the redshift dependence of the galaxy 2-point correlation function (2pCF), 
measured as a function of angle between the galaxy pair and line-of-sight (LoS).
As an illustration, Fig.\ref{fig_xy} shows how the shape distortion of objects (anisotropy of clustering signal) varies with distance (or redshift) when incorrect values of  $w_0$ or $w_a$ are used to infer distances.

\begin{figure*}
   \centering{
   \includegraphics[height=5cm]{fig_xy.eps}
   %\includegraphics[height=8cm]{fig4_1.eps}
   %\includegraphics[height=7cm]{FVol.eps}
   %\includegraphics[height=8cm]{xy.eps}
   %\includegraphics[height=8cm]{smu.eps}
   }
   \caption{\label{fig_xy}
   The distance dependence of the shape distortion in four wrongly assumed cosmologies,
   assuming a true cosmology of $\Omega_m=0.31$, $w=-1$.
   There are four perfect squares with their true shapes plotted in blue dashed lines.
   They are measured by an observer located at the origin, and the distances are reprojected in the four wrong cosmologies.
   The apparently distorted shapes are plotted in red solid lines.
   }
\end{figure*}

%$\xi_{\Delta s}(\mu)\equiv \int_{s_{\rm min}}^{s_{\rm max}} xi(s,\mu) ds$
%($s$ is the clustering scale).
%The RSD effect significantly distorts the shape of $\xi_{\Delta s}(mu)$ but exhibits much less redshift evolution 
%compared to the amount of change in due to incorrectly adopted cosmologies .

%When incorrect cosmological parameters are adopted, the shape of $\xi(\mu)$
%appears anisotropic due to the AP effect, and the amplitude
%is shifted by the change in comoving volume; both effects
%have significant redshift dependence. We test the method
%using the 2pCF on mock surveys drawn from HR3 and
%find the constraints obtained are tighter than those from
%the methodology of Li et al. (2014).

\citep{Li2016} applied their AP method to galaxies from the 3rd incarnation of the Sloan Digital Sky Survey (SDSS-III).
Combining their method with measurements of the Cosmic Microwave Background (CMB), type Ia supernovae (SNIa), 
baryon acoustic oscillations (BAO), and $H_0$,
they obtained very tight constraints of $ \Omega_m = 0.301 \pm 0.006,\ w=-1.054 \pm 0.025$.
In reducing the RSD effect, 
they were able to use galaxy clustering on scales down to 6 $h^{-1}$Mpc,
which is a major advance in extracting cosmological information 
on small scales where galaxy clustering is strong and there are many independent structures.

In this paper, we continue to develop our previous methodology and proceed to infer the constraints on dynamical dark energy. We will present a slightly improved methodology to that of \citep{Li2016}, 
allowing for faster likelihood estimation and thus the exploration of larger, higher dimensional parameter spaces. 
The methodology we will present here, can be applied to any model of dynamical dark energy, 
or indeed any appropriately chosen parametric or non-parametric decomposition of the cosmic expansion history. 
However, as a first step, in this paper we will focus on the widely used Chevallier-Polarski-Linder (CPL) parametrization
$w(z) = w_0 + w_a \frac{z}{1+z}$.

@@

Following our previous methodology \citep{Li2016}, the information of small-scale anisotropic clustering is computed
\footnote{These correlations were computed using the public code \texttt{KSTAT} \citep{kstat}.} 
as $\xi_{\Delta s} (\mu) \equiv \int_{s_{\rm min}}^{s_{\rm max}} \xi (s,\mu)\ ds$, 
with $s_{\rm min}=6 h^{-1} {\rm Mpc},\ s_{\rm max}=40 h^{-1} {\rm Mpc}$.
We then normalise these {\em clustering shells} as 
$\hat\xi_{\Delta s}(\mu) \equiv \frac{\xi_{\Delta s}(\mu)}{\int_{0}^{\mu_{\rm max}}\xi_{\Delta s}(\mu)\ d\mu}$
to nullify the amplitude information of the clustering signal, 
which is not associated with the Alcock-Paczynski test and is mostly sensitive to the galaxy bias.

@@

Systematics effects are estimated using mock catalogues derived from Horizon Run 4 (HR4) \cite{HR4},
an N-body simulation with box size $L={3150}$ $h^{-1}$Mpc, number of particles $6300^3$,   
initial redshift $z_{i}=100$, and WMAP5 cosmological parameters 
$(\Omega_{b},\Omega_{m},\Omega_\Lambda,h,\sigma_8,n_s)$  = (0.044, 0.26, 0.74, 0.72, 0.79, 0.96) \citep[]{komatsu 2011}. 
Mock galaxy samples are produced using a modified version of the one-to-one correspondence scheme \citep{hong2016}. 
The covariance matrix ${\bf Cov}$ is computed from the 2,000 sets of MultiDark PATCHY mock catalogues \citep{MDPATCHY}.
The statistical bias and scattering in the likelihood function (due to the finite number of mocks in covariance estimation) 
are corrected following \citet{Hartlap} and \citet{Percival2014}.

In \cite{Li2016} the likelihood contour of $\Omega_m-w$ was constructed by
measuring the 2PCF 3,375 times,
using 3D positions of BOSS galaxies computed in 71$\times$45 sets of cosmological parameters.
%with $0.06\leq \Omega_m\leq 0.41$, $-1.5 \leq w \leq -0.4$.
This procedure took 1 month using 500 cores of the KIAS {\texttt baekdu} cluster.
It would be too expensive to do such a computation on a 3D grid of $\Omega_m-w_0-w_a$,
so here we take an approach  of ``approximate 2PCF'', described as follows.

The number of pairs are only counted in a fiducial cosmology
-- which is taken as the $\Omega_m=0.26$ $\Lambda$CDM --
and ``translated'' to the measurements in a ``target'' cosmology using the relations, 
\begin{eqnarray}
 s_{\rm target} = s_{\rm fiducial} \sqrt{\alpha_{\parallel}^2 \mu_{\rm fiducial}+\alpha_{\bot}(1-\mu_{\rm fiducial}^2)}, \\
 \mu_{\rm target} = \mu_{\rm fiducial} \frac{\alpha_\parallel}
 {\sqrt{\alpha_{\parallel}^2 \mu_{\rm fiducial} +\alpha_{\bot}(1-\mu_{\rm fiducial}^2)}}
\end{eqnarray}
where $\alpha_{\bot}\equiv D_{A,\rm target}/D_{A,\rm fiducial}$,
$\alpha_{\parallel}\equiv H_{\rm fiducial}/H_{\rm target}$.
and values of $D_A$ and $H$ are computed in the effective redshifts of the six redshift bins.
In the fiducial cosmology
we measure $\xi(s,\mu)$ in an ultra high resolution of
$\Delta s = 0.2 {\rm Mpc/h}$, $\Delta \mu = 1/600$,
%Let us call this ``approximate 2pCF'' method.
%to obtain the accurate number counts 
and these small ``pixels'' are grouped to infer 
the number counts in other cosmologies, 
in large pixels of $\Delta s = 1 {\rm Mpc/h}$, $\Delta \mu = 1/120$.
%in the fiducial cosmology and group the bins in the target cosmology.
In the case when one small pixel belongs to more than one larger pixel (``edge effect'' ),
a correction is applied by computing the fraction of the overlapping area.
We found when using the approximate method the computed 
$\hat\xi_{\Delta s}(\mu)$ in a non-fiducial cosmology suffers from
an error of $\approx0.5\%$, which is 10 times smaller than 
the statistical noise in the 2pCF.

@@

\begin{figure*}
   \centering{
   %\includegraphics[width=9cm,natwidth=4,natheight=4]{figCPL_a.pdf}
   \includegraphics[height=5.3cm]{fig_nbar.eps}
   \includegraphics[width=7cm,]{fig_2pcf.eps}
   }
   \caption{\label{fig_TpCF}
   Left panel: We split the BOSS DR12 galaxies into six redshift bins to measure the evolution of shape distortion.
   The blue and green solid histograms show the redshift density distribution of LOWZ and CMASS galaxies in the $\Omega_m=0.31$ $\Lambda$CDM. 
   The vertical dashed lines define the six redshift bins that are used to cut the samples.
%   Their redshift ranges are listed.                                       .
   The number of LOWZ/CMASS galaxies in the three low/high redshift bins are listed.
   Right panel: 2D contour map of measured $\xi$ as a function of $\mu$ and $s$, from the six redshift bins of LOWZ and CMASS samples, 
      in the cosmology of $\Omega_m=0.31$ $\Lambda$CDM model.
    The black dashed lines mark the clustering region 6-40\ Mpc/h studied in the AP method.
    %The correlation functions have rather large values on small sales such as $s=5$ $h^{-1}$Mpc,
    %  and monotonically decreases to smaller values on larger scales due to weaker clustering.
    The six contour maps have rather similar appearance, implying small redshift evolution of $\xi$ from RSD.
   }
\end{figure*}

The Planck team has released the {\texttt {COSMOMC}} \citep{LB2002} outputs of four MCMC ``chains'' in the CPL model, 
using a combination of four datasets:
the full-mission Planck observations of CMB temperature and polarization anisotropies \cite{Planck2015},
the BAO distance priors measured from SDSS DR11 \citep{Anderson2013}, 6dFGS \citep{6dFGS} and SDSS MGS \citep{MGS},
the ``JLA'' SNIa sample \citep{JLA},
and the Hubble Space Telescope measurement of $H_0=70.6\pm3.3$ \cite{Riess2011,E14H0}.
\footnote{http://pla.esac.esa.int/pla/\#cosmology}.
These MCMC chains contain the CMB+BAO+SNIa+$H_0$ likelihood computed for $\sim$37,000 sets of cosmological parameters,
yielding constraints of 
\begin{equation}
 \Omega_m = 0.309 \pm 0.010, w_0 = -0.938 \pm 0.109, w_a = -0.38 \pm 0.41.
\end{equation}
%To include the AP method into the constraints,
We then compute the likelihood of the LOWZ+CMASS AP measurements for each of the parameter combinations provided in the Planck chain. After multiplying the sample likelihoods with ours, while also changing the sample weights {\bf (HOW CHANGE?)}, we derive the CMB+BAO+SNIa+$H_0$+AP constraints on CPL parameters,
\begin{equation}
%\Omega_m = 0.300 \pm 0.008, w_0 = -1.056 \pm 0.061, w_a = -0.04 \pm 0.27,
    %1  0.2999656E+00  0.7616334E-02  0.2958745E+00  0.3035033E+00  0.2878169E+00  0.3132870E+00   \Omega_m
    %2  0.6904801E+00  0.8675444E-02  0.6808558E+00  0.6993898E+00  0.6722945E+00  0.7052207E+00   h
    %3 -0.1056468E+01  0.6096323E-01 -0.1116874E+01 -0.9965903E+00 -0.1175041E+01 -0.9287388E+00   w
    %4 -0.3808376E-01  0.2716607E+00 -0.3105235E+00  0.2321461E+00 -0.6033103E+00  0.4782076E+00   w_a    
\Omega_m = 0.301 \pm 0.008, w_0 = -1.042 \pm 0.067, w_a = -0.07 \pm 0.29.
%    1  0.3014353E+00  0.7759923E-02  0.2974325E+00  0.3048882E+00  0.2889838E+00  0.3152269E+00   \Omega_m
%    2  0.6887652E+00  0.8760036E-02  0.6797134E+00  0.6978279E+00  0.6703971E+00  0.7044224E+00   h
%    3 -0.1041626E+01  0.6732817E-01 -0.1107730E+01 -0.9768319E+00 -0.1172149E+01 -0.8979017E+00   w
%    4 -0.7173374E-01  0.2860547E+00 -0.3572176E+00  0.2107347E+00 -0.6775045E+00  0.4558817E+00   w_a    
\end{equation}
%$$.
While the central values of parameters do not shift appreciably,
the error bars of $w_0$ and $w_a$ are reduced by 38.4\% and 29.7\%, respectively.% $$-40\%$ by adding our method into the analysis.

@@

The derived cosmological constraints are fully consistent with a cosmological constant dark energy component having no evolution.
Adding the new AP method to the CMB+BAO+SNIa+$H_0$ constraints reduces the statistical errors by $\sim$30-40\%,
and reduces the 95.4\% CL $w_0-w_a$ contour by 50\%.
This reveals the constraining power of the small scale AP effect for probing the redshift evolution of dark energy.

%\cite{Li2016} shows that the AP method outperforms SNIa in constraining the parameter space of $\Omega_m-w$.
%Similarly we find it outperforms in constraining $w_0-w_a$, as shown in the middle panel of Fig.\ref{fig_TpCF}.

\begin{figure*}
   \centering{
   %\includegraphics[width=9cm,natwidth=4,natheight=4]{figCPL_a.pdf}
   \includegraphics[height=5cm,natwidth=6,natheight=4.5]{figCPL_b.pdf}
   %\includegraphics[width=5.5cm,natwidth=4,natheight=4]{figCPL_b_BAOSNIaAP.pdf}
   \includegraphics[height=5cm,natwidth=6,natheight=4.5]{figCPL_d.pdf}
   %\includegraphics[height=8cm]{fig2b.pdf}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
   }
   \caption{\label{fig_con}
   Left panel: 68.3\%, 95.4\% CL likelihood contours in the  $w_0-w_a$ plane.
   %Results are consistent with $\Lambda$CDM.
   The constrained area is significantly reduced after combined with the AP method.
   %This shows the power of the AP method in constraining dynamical dark energy.
   %Middle panel: $w_0-w_a$ likelihood contours using datasets of CMB+BAO, CMB+BAO+SNIa+$H_0$ and CMB+BAO+AP+$H_0$.
   %When combined with CMB+BAO, AP method yields much tighter constraints than SNIa.
   Right panel: Redshift evolution of $w(z)$, with/without adding the AP constraint.
   Adding AP tightens the constraints and reduces the redshift evolution of $w$ (tilt of $w(z)$).
   }
\end{figure*}

In this short paper we have chosen to focus on the popular CPL parameterisation of dark energy, 
however our methodology could be used with any model of cosmic expansion history.

%%% To be written!!!
%\section{Robustness Test}\label{sec:test}
{\it Robustness Test.}---

Analysing the same DR12 BOSS data, \citet{Li2016} obtained constraints on $\Omega_m$ and $w$
\begin{equation}\label{eq:wcdm_constrain_default}
\Omega_m=0.301\pm 0.006, w=-1.054\pm 0.025.
\end{equation}
In our new ``approximate 2pCF'' approach, while taking a fiducial cosmological model of $\Omega_m=0.26, w=-1$, we obtain consistent results of,
\begin{equation}
\Omega_m = 0.301 \pm 0.008,\ w=-1.053\pm 0.034.
%    1  0.3010870E+00  0.7818562E-02  0.2973018E+00  0.3045011E+00  0.2883318E+00  0.3150334E+00   \Omega_m
%    2  0.6886453E+00  0.8507156E-02  0.6802557E+00  0.6975813E+00  0.6708444E+00  0.7047652E+00   h
%    3 -0.1053343E+01  0.3395215E-01 -0.1087474E+01 -0.1020395E+01 -0.1121890E+01 -0.9819443E+00   w
\end{equation}
Since we use a smaller $n_{\mu}$ value than that adopted by \citep{Li2016}
\footnote{\citep{Li2016} adopted $n_{\mu}=25-35$ to achieve tighter constraints.
Since we are using an approximate 2pCF, for robustness considerations we use $n_{\mu}=20-25$.}, 
resulting in a slightly enlarged error bar. 
The central values of the parameters agree quite well, demonstrating the reliability of the approximate 2pCF method.

%\begin{figure*}
%   \centering{
%   \includegraphics[width=6cm]{figIO.eps}
%   %\includegraphics[width=9cm,natwidth=4,natheight=4]{fig2b.pdf}
%   %\includegraphics[height=8cm]{fig2b.pdf}
%   %   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%   %    \includegraphics[height=8cm]{smu.eps}
%   }
%   \caption{\label{fig_IO}
%   Input-output test of the AP method. 
%   The input cosmology (marker by green plus) is recovered at $\sim$0.5$\sigma$ 
%   using 4 realizations.
%   }
%\end{figure*}

>>>>

\author{ 
Xiao-Dong Li,$^{1}$
Cristiano G. Sabiu,$^{2}$
Changbom Park,$^{1}$
Yuting Wang,$^{3}$
Gong-bo Zhao,$^{3}$
Hyunbae Park,$^{2}$
Arman Shafieloo,$^{2}$
Juhan Kim,$^{4,1}$
and Sungwook E. Hong$^{2}$}

@@

\begin{abstract}
We use the small scale anisotropic clustering signal of 1,133,326 galaxies from the Sloan Digital Sky Survey Data Release 
12, spanning the redshift range $0.15<z<0.69$ to place constraints on the expansion history of the Universe. 
Concentrating on the Chevallier-Polarski-Linder parametrization of dark energy, $w=w_0+w_a\frac{z}{1+z}$,
while improving upon earlier works \citep{Li2016}, 
we present a novel use of the Alcock-Paczynski (AP) test via the small scales clustering difference between redshift bins. 
This differential measure is shown to be a robust statistical tool that mitigates many systematic effects, 
allowing for competitive and unbiased constrains to be placed on cosmological parameters.
We find that our results, in combination with the CMB, BAO, SNIa and $H_0$ from Cepheid data yield, 
$\Omega_m = 0.301 \pm 0.008,\ w_0 = -1.042 \pm 0.067,\ w_a = -0.07 \pm 0.29$ (68.3\% CL). 
Adding our new AP measurements to the aforementioned results reduced the $w_0-w_a$ 2-$\sigma$ contour area by 50\%.
The AP method can be combined with the standard BAO method, 
greatly increasing the power of ongoing and future galaxy surveys in probing dark energy.
\end{abstract}

@@

%\section{Introduction}
{\it Introduction.}---
The  origin  of  the late-time accelerating  expansion  of  the  universe  is  one of  the  most  salient  questions  in  contemporary cosmology. Theoretical explanations for this phenomena are many and range from a non-zero  vacuum energy and evolving  scalar  fields  remnant  from  the  big  bang, to modifications of Einstein's General Relativity\cite{2012IJMPD..2130002Y}. 
Considering the wealth of theoretical explanations, it is crucial to obtain precise and unbiased measurements of the expansion history of the Universe allowing us to 
differentiate between competing models.

@@

Previously, we proposed to use the  {\it redshift dependence} of the AP distortion\cite{Li2014} as a way of mitigating the RSD effect. The clustering anisotropies produced by RSD are, although large, close to uniform in magnitude over a wide range in redshift.  However, if cosmological parameters are incorrectly chosen resulting in an AP effect, 
the anisotropy in the clustering signal has a clear redshift dependence
(as an illustration, Fig.\ref{fig_xy} shows how the shape distortion of objects varies with distance (i.e., redshift) 
when incorrect cosmologies are used to infer distances).
In a later work\cite{Li2015}, we developed an AP methodology 
that utilises the redshift dependence of the galaxy 2-point correlation function (2pCF), 
measured as a function of angle between the galaxy pair and line-of-sight (LoS).

In an earlier work\citep{Li2016} we applied this AP method to galaxies from the 3rd incarnation of the Sloan Digital Sky Survey (SDSS-III). 
Combining the method with measurements of the Cosmic Microwave Background (CMB), type Ia supernovae (SNIa), 
baryon acoustic oscillations (BAO), and $H_0$,
we obtained very tight constraints of $ \Omega_m = 0.301 \pm 0.006,\ w=-1.054 \pm 0.025$.
In reducing the RSD effect, 
we were able to use galaxy clustering on scales down to 6 $h^{-1}$Mpc,
which is a major advance in extracting cosmological information 
on small scales where galaxy clustering is strong and there are many independent structures.

\begin{figure*}
   \centering{
   \includegraphics[height=5cm]{fig_xy.eps}
   %\includegraphics[height=8cm]{fig4_1.eps}
   %\includegraphics[height=7cm]{FVol.eps}
   %\includegraphics[height=8cm]{xy.eps}
   %\includegraphics[height=8cm]{smu.eps}
   }
   \caption{\label{fig_xy}
   The distance dependence of the shape distortion in four wrongly assumed cosmologies.
   Assuming a true cosmology of $\Omega_m=0.31$, $w=-1$.
   There are four perfect squares with their true shapes plotted in blue dashed lines.
   They are measured by an observer located at the origin, and the distances are reprojected in the four wrong cosmologies.
   The apparently distorted shapes are plotted in red solid lines.
   }
\end{figure*}

In this paper, we continue to develop our previous methodology and proceed to infer the constraints on dynamical dark energy. 
We will present a slightly improved methodology to that of our previous work\citep{Li2016}, 
allowing for faster likelihood estimation and thus the exploration of larger, higher dimensional parameter spaces. 
The methodology we will present here, can be applied to any model of dynamical dark energy, 
or indeed any appropriately chosen parametric or non-parametric decomposition of the cosmic expansion history. 
However, as a first step, in this paper we will focus on the widely used Chevallier-Polarski-Linder (CPL) parametrization\cite{CPL_CP,CPL_L},
\begin{equation}
w(z) = w_0 + w_a (1-a) = w_0 + w_a \frac{z}{1+z}.
\end{equation}

@@

Following our previous methodology \citep{Li2016}, the information of small-scale anisotropic clustering is computed.
%\footnote{These correlations were computed using the public code \texttt{KSTAT} \citep{kstat}.} 
as $\xi_{\Delta s} (\mu) \equiv \int_{s_{\rm min}}^{s_{\rm max}} \xi (s,\mu)\ ds$, 
with $s_{\rm min}=6 h^{-1} {\rm Mpc},\ s_{\rm max}=40 h^{-1} {\rm Mpc}$.
We then normalise these {\em clustering shells} as 
$\hat\xi_{\Delta s}(\mu) \equiv \frac{\xi_{\Delta s}(\mu)}{\int_{0}^{\mu_{\rm max}}\xi_{\Delta s}(\mu)\ d\mu}$
to nullify the amplitude information of the clustering signal, 
which is not associated with the Alcock-Paczynski test and is mostly sensitive to the galaxy bias.

@@

Systematics effects are estimated using mock catalogues derived from Horizon Run 4 (HR4) \cite{HR4},
an N-body simulation with box size $L={3150}$ $h^{-1}$Mpc, number of particles $6300^3$,   
initial redshift $z_{i}=100$, and WMAP5\citep{komatsu2011} cosmological parameters 
$(\Omega_{b},\Omega_{m},\Omega_\Lambda,h,\sigma_8,n_s)$  = (0.044, 0.26, 0.74, 0.72, 0.79, 0.96). 
Mock galaxy samples are produced using a modified version of the one-to-one correspondence scheme \citep{hong2016}. 
The covariance matrix, ${\bf Cov}$, is computed from the 2,000 sets of MultiDark PATCHY mock catalogues \citep{MDPATCHY}.
The statistical bias and scattering in the likelihood function (due to the finite number of mocks in covariance estimation) 
are adequately corrected\citet{Hartlap,Percival2014}.

In our earlier work\cite{Li2016}, the likelihood contour of $\Omega_m-w$ was constructed by
measuring the 2PCF 3,375 times,
using 3D positions of BOSS galaxies computed in 71$\times$45 sets of cosmological parameters.
%with $0.06\leq \Omega_m\leq 0.41$, $-1.5 \leq w \leq -0.4$.
This procedure took 1 month using 500 cores of the KIAS {\texttt {baekdu}} cluster.
It would be computationally intractable to attempt a full MCMC of all relevant cosmological parameters using this approach. 
Thus we adopt an ``approximate 2PCF'' by transforming our measurements from one cosmology to another.
%It would be too expensive to do such a computation on a 3D grid of $\Omega_m-w_0-w_a$,
%so here we take an approach  of ``approximate 2PCF'', described as follows.

In detail, the number of pairs are only counted in a fiducial cosmology
-- which is taken as the $\Omega_m=0.26$ $\Lambda$CDM --
and ``translated'' to the measurements in a ``target'' cosmology using the relations, 
\begin{eqnarray}
 s_{\rm target} = s_{\rm fiducial} \sqrt{\alpha_{\parallel}^2 \mu_{\rm fiducial}+\alpha_{\bot}(1-\mu_{\rm fiducial}^2)}, \\
 \mu_{\rm target} = \mu_{\rm fiducial} \frac{\alpha_\parallel}
 {\sqrt{\alpha_{\parallel}^2 \mu_{\rm fiducial} +\alpha_{\bot}(1-\mu_{\rm fiducial}^2)}}
\end{eqnarray}
where $\alpha_{\bot}\equiv D_{A,\rm target}/D_{A,\rm fiducial}$,
$\alpha_{\parallel}\equiv H_{\rm fiducial}/H_{\rm target}$.
and values of $D_A$ and $H$ are computed in the effective redshifts of the six redshift bins.
In the fiducial cosmology
we measure $\xi(s,\mu)$ in an ultra high resolution of
$\Delta s = 0.2 {\rm Mpc/h}$, $\Delta \mu = 1/600$,
%Let us call this ``approximate 2pCF'' method.
%to obtain the accurate number counts 
and these small ``pixels'' are grouped to infer 
the number counts in other cosmologies, 
in large pixels of $\Delta s = 1 {\rm Mpc/h}$, $\Delta \mu = 1/120$.
%in the fiducial cosmology and group the bins in the target cosmology.
In the case when one small pixel belongs to more than one larger pixel,
a correction is applied by computing the fraction of the overlapping area.

We found when using the approximate method the computed 
$\hat\xi_{\Delta s}(\mu)$ in a non-fiducial cosmology suffers from
an error of $\approx0.5\%$, which is 10 times smaller than 
the statistical noise in the 2pCF.
As a result, the approximate method reproduces the results presented in \cite{Li2016} quite well.

@@

\begin{figure*}
   \centering{
   %\includegraphics[width=9cm,natwidth=4,natheight=4]{figCPL_a.pdf}
   \includegraphics[height=5.3cm]{fig_nbar.eps}
   \includegraphics[width=7cm,]{fig_2pcf.eps}
   }
   \caption{\label{fig_TpCF}
   Spliting the SDSS galaxies into six redshiftbins, and measuring the correlation function.
   Left panel: We split the BOSS DR12 galaxies into six redshift bins to measure the evolution of shape distortion.
   The blue and green solid histograms show the redshift density distribution of LOWZ and CMASS galaxies in the $\Omega_m=0.31$ $\Lambda$CDM. 
   The vertical dashed lines define the six redshift bins that are used to cut the samples.                                       .
   The number of LOWZ/CMASS galaxies in the three low/high redshift bins are listed.
   Right panel: 2D contour map of measured $\xi$ as a function of $\mu$ and $s$, from the six redshift bins of LOWZ and CMASS samples, 
      in the cosmology of $\Omega_m=0.31$ $\Lambda$CDM model.
    The black dashed lines mark the clustering region 6-40\ Mpc/h studied in the AP method.
    The six contour maps have rather similar appearance, implying small redshift evolution of $\xi$ from RSD.
   }
\end{figure*}

The CPL parametrization characterizes dark energy equation-of-state (EoS) by two free parameters.
$w_0$ determines the present-day value, while parameter $w_a$ characterizes the first-order derivative of $w$ with respect to $a$. 
The CPL parametrization has many obvious advantages, for instance, a manageable parameter space, 
the bounded behavior at high redshift, 
and the ability to accurately reconstruct many dark energy theories\cite{CPL_L}. 
So it was widely used in the literature to reconstruct dark energy from observations.
The Planck team has released the {\texttt {COSMOMC}} \citep{LB2002} outputs of four MCMC ``chains'' in the CPL model, 
using a combination of four datasets:
the full-mission Planck observations of CMB temperature and polarization anisotropies \cite{Planck2015},
the BAO distance priors measured from SDSS DR11 \citep{Anderson2013}, 6dFGS \citep{6dFGS} and SDSS MGS \citep{MGS},
the ``JLA'' SNIa sample \citep{JLA},
and the Hubble Space Telescope measurement of $H_0=70.6\pm3.3$ \cite{Riess2011,E14H0}.
These MCMC chains contain the CMB+BAO+SNIa+$H_0$ likelihood computed for $\sim$37,000 sets of cosmological parameters,
yielding constraints of %$ \Omega_m = 0.309 \pm 0.010, w_0 = -0.938 \pm 0.109, w_a = -0.38 \pm 0.41$.
\begin{equation}
 \Omega_m = 0.309 \pm 0.010, w_0 = -0.938 \pm 0.109, w_a = -0.38 \pm 0.41.
\end{equation}
%To derive the constraints combined with our AP method,
%we then compute the likelihood of the SDSS LOWZ+CMASS AP measurements for each of the parameter combinations provided in the Planck chain. 
%To do this, we splited the 1,133,326 galaxies from the SDSS DR12 into six redshift bins,
%measured the 2pCF (on scales of 6-40 $h^{-1}$Mpc) in these redshift bins (see Fig.\ref{fig_nbar}).
%This enables us investigating the redshift evolution of anisotropic clustering.
%We subtracted the systematic effects based on estimations from mocks made from Horizon Run 4 simulations \citep{HR4},
%and used a $\chi^2$ function to assess the redshift dependence.
%(see Methods for further explanations).
After adding the log-likelihoods of the Planck team sample with ours, 
while also multiply the sample weights by our likelihoods, 
we derive the CMB+BAO+SNIa+$H_0$+AP constraints on CPL parameters, %$\Omega_m = 0.301 \pm 0.008, w_0 = -1.042 \pm 0.067, w_a = -0.07 \pm 0.29$.
\begin{equation}
\Omega_m = 0.301 \pm 0.008, w_0 = -1.042 \pm 0.067, w_a = -0.07 \pm 0.29.
\end{equation}
While the central values of parameters do not shift appreciably,
the error bars of $w_0$ and $w_a$ are reduced by 38.4\% and 29.7\%, respectively.% $$-40\%$ by adding our method into the analysis.

@@

The derived cosmological constraints are fully consistent with a cosmological constant dark energy component having no evolution.
Adding the new AP method to the CMB+BAO+SNIa+$H_0$ constraints reduces the statistical errors on many parameters by $\sim$30-40\% 
and reduces the $w_0-w_a$ (95.4\% CL) contour by 50\%.
This reveals the constraining power of the small scale AP effect for probing the redshift evolution of dark energy.

We performed a thorough test of all the possible variables in our methodology.
Figure \ref{fig_contest} shows that 
the result is unaffected by the line-of-sight $\mu$-cut, 
the number of angular bins, 
the range of radial integration and the choice of fiducial cosmology. 
%we adopt $\mu_{\rm max}=0.99\ (0.85)$ to include more (less) regions near the LOS,
%decrease the number of angular bins to $n_{\mu}=15-20$,
%change the range of clustering to $8-30 \rm Mpc/h$,
%change the fiducial cosmology to $\Omega_m=0.26,\ w=-0.6$,
%exclude 1 or 2 high redshift bins,
%or reduce the number of mocks to 1000.
The most striking point to note in our testing is that the 
shift in contour position is very small even when we do not include the systematic correction.
This is reassuring, since we may have worried that {\it cosmological dependence} 
of the systematic correction could influence the final result.

\begin{figure*}
   \centering{
   %\includegraphics[width=9cm,natwidth=4,natheight=4]{figCPL_a.pdf}
   \includegraphics[height=5cm,natwidth=6,natheight=4.5]{figCPL_b.pdf}
   %\includegraphics[width=5.5cm,natwidth=4,natheight=4]{figCPL_b_BAOSNIaAP.pdf}
   \includegraphics[height=5cm,natwidth=6,natheight=4.5]{figCPL_d.pdf}
   %\includegraphics[height=8cm]{fig2b.pdf}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
   }
   \caption{\label{fig_con}
   Cosmological constraints on the CPL dark energy parametrization $w=w_0+w_a\frac{z}{1+z}$.
   Left panel: 68.3\%, 95.4\% CL likelihood contours in the  $w_0-w_a$ plane.
   %Results are consistent with $\Lambda$CDM.
   The constrained area is significantly reduced after combined with the AP method.
   %This shows the power of the AP method in constraining dynamical dark energy.
   %Middle panel: $w_0-w_a$ likelihood contours using datasets of CMB+BAO, CMB+BAO+SNIa+$H_0$ and CMB+BAO+AP+$H_0$.
   %When combined with CMB+BAO, AP method yields much tighter constraints than SNIa.
   Right panel: Redshift evolution of $w(z)$, with/without adding the AP constraint.
   Adding AP tightens the constraints and reduces the redshift evolution of $w$ (tilt of $w(z)$).
   }
\end{figure*}

%Covariance matrix

\begin{figure*}
   \centering{
   \includegraphics[width=16cm,natwidth=6,natheight=5]{fig_con_tests.pdf}
   %\includegraphics[width=9cm,natwidth=4,natheight=4]{fig2b.pdf}
   %\includegraphics[height=8cm]{fig2b.pdf}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
   }
   \caption{\label{fig_contest}
   Robustness test of the results.
   For the ``default'' constraints (pink contours) are derived using six redshift bins, 
   $\mu_{\rm max}=0.97$, $n_{\mu}=15-20$, $s=6-40 \rm Mpc/h$,
   a fiducial cosmology $\Omega_m=0.26,\ w=-1.0$ for the measurement of 2pCF,
   systematic effects estimated from HR4 simulations, 
   and covariance estimated using 2,000 PATCHY mocks.
   When altering these options, the results remain robust (black contours).
   }
\end{figure*}
